### 2015-11-16 ch02-qlogms/excitation-design.tex
Quite some changes to CH02

\section{Introduction}
\label{sec:excitation:intro}
Experiment design is the first step in the system identification loop, as illustrated in \figref{fig:intro:identification-cycle}.
Consequently, the design that is chosen influences every further step in the identification process and limit the model quality that can be attained.
In many cases

\subsection{Generic excitation signals}
\label{sec:excitation:intro:generic}

@@

@@

\subsection{Robust excitation signals}
\label{sec:excitation:intro:robust}
As illustrated above, optimal experiment design places the user in a catch-22 situation.

\subsection{Approach in this chapter}
\label{sec:excitation:intro:approach}
In this chapter, the goal is to construct excitation signals that bridge the properties of generic and robust excitation signals to alleviate the chicken-and-the-egg problem that optimal strategies typically incur.
In particular, a generic excitation signal will be used but its properties are tuned in such a way that it

@@

  \subsection{The system under consideration}
  We consider the transfer function models of generic \gls{LTI} systems of arbitrary but finite orders $n_a\times n_b$:
  \begin{align}
    G\left( \Omega, \theta \right) = \frac{B\left( \Omega,\theta \right)}
                                          {A\left( \Omega,\theta \right)}
                                   = \frac{\sum\limits_{i=0}^{n_b-1} b_i \Omega^i}
                                          {\sum\limits_{i=0}^{n_a-1} a_i \Omega^i}
    \text{.} \label{eq:model}
  \end{align}

@@

  Such a system can be decomposed into a series of first and second order systems by the partial fraction expansion.

  To study the behavior of one sub-system, we inspect a generic second order system $G\left( s \right)$.
  In continuous time such a system can be described by a transfer function of the form
  \begin{equation}
    G\left(s \right) = \frac{\dc}
                            {             \frac{s^2}
                                               {\wn^2} 
                             + 2 \damping \frac{s}
                                               {\wn}
                             + 1
                            }
  \text{.}
  \label{eq:secondOrderSystemCT}
  \end{equation}
\TODO{This is actually wrong, rather, it should be $\frac{as-b}{cs^2 + ds + 1}$. However, the difference in information is rather small around the actual resonance. See script in NnF folder.}
\TODO{write out in partial fraction expansion, use script to show that in practice, only a few parameters matter. This will be well-approximated by considering only the second order system of the paper.}

@@

\subsection{System and noise model}
  In this paper, an output error (OE) system setup is used.
  Denote
    $U  \left(\omega\right)$ as the excitation signal,
    $Y_0\left(\omega\right)$ the exact, unperturbed system output,
    $E  \left(\omega\right)$ for the output noise source,
    $Y  \left(\omega\right) = Y_0\left( \omega \right) + E\left( \omega \right)$ depicts the measured output signal.

@@

  Consequently, the complete multisine has a period $T_{u(t)}$ that is the common multiple of the period of all its components:
  \begin{equation}
    T_{u(t)}  = \lcm \set{T_1, \ldots, T_F}
      = \lcm \set{f_1^{-1}, \ldots, f_F^{-1}}
    \text{.}
  \end{equation}
  \begin{definition}[Least common (integer) multiple] \label{def:lcm-real}
  The least common multiple of the set of values $T_k \in \RealNumbers$ for $k \in \set{1, \ldots, F}$ is  
  \begin{equation}
    \lcm \set{T_1, \ldots, T_F} \isdef
      \min \set{T \given{\forall k \in \set{1,\ldots,F}, \exists n_k \in \NaturalNumbers: n_k T_k = T }}
   \end{equation}
   if such a minimum exists.
   \end{definition}
  \begin{example}
   Consider a generalized multisine where $f_1 = 1 \unit{Hz}$ and $f_2 = \tfrac{2}{3} \unit{Hz}$.
   The corresponding periods are $T_1 = f_1^{-1} = 1 \unit{s}$ and $T_2 = f_2^{-1} = 1.5 \unit{s}$.
   The overall period of the multisine is $T = n_1 T_1 = n_2 T_2 = 3 \unit{s}$, i.e. $n_1 = 3$ and $n_2=2$.
  \end{example}

@@

  \TODO{link naar log meas mode DSA (e.g. HP3562A) leggen. Bij HP 3562A, is $F/\mathrm{decade} = 80$, oftewel $\alpha = \sqrt[80]{10} \approx 1.029$ vast ingesteld. Dit beperkt bruikbaarheid van zo'n toestel.}

  \begin{property}
   Each frequency band $B = \ClosedInterval{f, \kappa f}$ with $\kappa \in \PositiveRealNumbers \without \set{0,1}$ that is completely within the range of the excited logarithmic grid of frequencies $\set{f_1, \ldots, f_F}$ contains a constant number of frequency lines
   \begin{equation}
     F_{B}(f) \isdef
     \floor{\frac{\log (\kappa f) - \log( f )}{\log \alpha}}
         = \floor{\log_{\alpha} \kappa}
         \text{.}
   \end{equation}
   \end{property}
  \begin{example}
    For the frequency decade band $\decade(f) \isdef \ClosedInterval{f, 10 f}$, a logarithmic frequency grid contains $F_{\decade} = \floor{\log_{\alpha} 10}$ excited lines.
  \end{example}

@@

  Using this extension, the frequency range covered by $f_k$ can then be calculated as
    \begin{align}
      \fMin{k} &= \mean{f_{k-1}}{f_k}\\
      \fMax{k} &= \mean{f_{k+1}}{f_k}
    \label{eq:fRange}
    \text{.}
    \end{align}
  This range can be seen as the power spectral density counterpart of the discrete spectrum relation \eqref{eq:nearestLines}.
  Note that instead of the geometric mean, the arithmetic mean can be used in \eqref{eq:fRange} to calculate almost identical boundaries.

@@

  The obtained FRF is then fitted to the transfer function model \eqref{eq:secondOrderSystemDTidentifiable} using the \texttt{elis} function of the \FDIDENT \citep{FDIDENT} for \matlab.
  This is a maximum likelihood estimator that fits a parametric model to the measured data in the frequency domain.
  To aid convergence and avoid local minima, the initial values for the model are chosen equal to the parameters of the true system $G_0(z)$, so that we can focus completely on the variance of the estimations.

@@

@@

  The resulting frequency response function $\hat{G}$ for one such system is shown in \figref{fig:FRF1} together with the standard deviation $\sigma_{\hat{G}}$ of the estimated parametric model for each of the different excitation signals.
  The bias is at least $20\unit{dB}$ smaller than the standard  deviation.
  Therefore the mean squared error (MSE) will be determined mainly by the variance of the estimate instead of the bias.

  \begin{figure}[th]
    \centering
    \setlength\figureheight{5cm}
    \setlength\figurewidth{0.8\columnwidth}
    \input{\thisDir/figs/Sys02-4.tikz}
    \caption[Simulated FRF and its variance for different excitation signals.]{FRF of a second order system with $\fR=40\unit{Hz}$ and $\damping=0.2$
             along with the corresponding variance.}%
    \label{fig:FRF1}
  \end{figure}

@@

\begin{subappendices}
  \section{Justification for considering only poles}

\end{subappendices}

>>>>

\section{Introduction}
\label{sec:excitation:intro}
Experiment design is the first step in the system identification loop, as illustrated in \figref{fig:intro:identification-cycle}.
Consequently, the design that is chosen influences every further step in the identification process and limit the model quality that can be attained.

\subsection{Generic excitation signals}
\label{sec:excitation:intro:generic}
%TODO

@@

%TODO

@@

\subsection{Robust excitation signals}
\label{sec:excitation:intro:robust}
As illustrated above, optimal experiment design places the user in a catch-22 situation.
In the recent years, this has sparked attention towards robust signal design~\citep{Rojas2007} to alleviate the drawbacks of (nominal) optimal experiment design.
%TODO

\subsection{Approach in this chapter}
\label{sec:excitation:intro:approach}
In this chapter, the goal is to construct excitation signals that bridge the properties of generic and robust excitation signals to alleviate the chicken-and-the-egg problem that optimal strategies typically incur.
In particular, a generic excitation signal will be used but its properties are tuned in such a way that it attains some properties of robust excitation signals without the requirement of very specific prior information.

@@

  \subsection{The system under consideration}
  We consider the transfer function models of generic \gls{LTI} systems of arbitrary but finite orders $n_b / n_a$:
  \begin{align}
    G\left( \Omega, \theta \right) = \frac{B\left( \Omega,\theta \right)}
                                          {A\left( \Omega,\theta \right)}
                                   = \frac{\sum\limits_{i=0}^{n_b-1} b_i \Omega^i}
                                          {\sum\limits_{i=0}^{n_a-1} a_i \Omega^i}
    \text{.} \label{eq:model}
  \end{align}

@@

  Such a system can be decomposed into a series of first and second order systems by the partial fraction expansion.
  \TODO{This is actually wrong, rather, it should be $\frac{as-b}{cs^2 + ds + 1}$. However, the difference in information is rather small around the actual resonance. See script in NnF folder.}
\TODO{write out in partial fraction expansion, use script to show that in practice, only a few parameters matter. This will be well-approximated by considering only the second order system of the paper.}

  Within these sub-systems, we can distinguish first-order subsystems and second-order subsystems, the latter of which describe resonant phenomena.
  Since such resonances are often a limiting factor, we first focus on these resonant sub-systems.
  Afterwards, it will become obvious that

  To study the behavior of one sub-system, we inspect a generic second order system $G\left( s \right)$.
  In continuous time such a system can be described by a transfer function of the form
  \begin{equation}
    G\left(s \right) = \frac{\dc}
                            {             \frac{s^2}
                                               {\wn^2} 
                             + 2 \damping \frac{s}
                                               {\wn}
                             + 1
                            }
  \text{.}
  \label{eq:secondOrderSystemCT}
  \end{equation}

@@

\subsection{System and noise model}
  In this chapter, an output error (OE) system setup is used.
  Denote
    $U  \left(\omega\right)$ as the excitation signal,
    $Y_0\left(\omega\right)$ the exact, unperturbed system output,
    $E  \left(\omega\right)$ for the output noise source,
    $Y  \left(\omega\right) = Y_0\left( \omega \right) + E\left( \omega \right)$ depicts the measured output signal as shown in \figref{fig:excitation:OE}.

  \begin{figure}
    \centering
    \input{\thisDir/figs/oesetup.tikz}
    \caption[Output-error]{Output-error set-up.}
  \end{figure}

@@

  Consequently, the complete multisine has a period $T_{u(t)}$ that is the common multiple of the period of all its components:
  \begin{equation}
    T_{u(t)}  = \lcm \set{T_1, \ldots, T_F}
      = \lcm \set{f_1^{-1}, \ldots, f_F^{-1}}
    \text{.}
  \end{equation}
  \begin{definition}[Least common (integer) multiple] \label{def:lcm-real}
  The least common multiple of the set of values $T_k \in \RealNumbers$ for $k \in \set{1, \ldots, F}$ is  
  \begin{equation}
    \lcm \set{T_1, \ldots, T_F} \isdef
      \min \Set{T |\forall k \in \set{1,\ldots,F}, \exists n_k \in \NaturalNumbers: n_k T_k = T }
   \end{equation}
   if such a minimum exists.
   \end{definition}
  \begin{example}
   Consider a generalized multisine where $f_1 = 1 \unit{Hz}$ and $f_2 = \tfrac{2}{3} \unit{Hz}$.
   The corresponding periods are $T_1 = f_1^{-1} = 1 \unit{s}$ and $T_2 = f_2^{-1} = 1.5 \unit{s}$.
   The overall period of the multisine is $T = n_1 T_1 = n_2 T_2 = 3 \unit{s}$, i.e. $n_1 = 3$ and $n_2=2$.
  \end{example}

@@

  \begin{remark}
    In different contexts, e.g. electrical circuit simulators such as \gls{SPICE}~\citep{Kundert1995}, measurement equipment such as \glspl{DSA} or the \code{logspace} function in \MATLAB it is more common to specify the spacing of a logarithmic grid using the total number of frequencies ($F$) or the number of frequencies per decade ($F_{\decade}$).
  \end{remark}

  \begin{property}
   Each frequency band $B = \ClosedInterval{f, \kappa f}$ with $\kappa \in \PositiveRealNumbers \without \set{0,1}$ that is completely within the range of the excited logarithmic grid of frequencies $\set{f_1, \ldots, f_F}$ contains a constant number of frequency lines
   \begin{equation}
     F_{B}(f) \isdef
     \floor{\frac{\log (\kappa f) - \log( f )}{\log \alpha}}
         = \floor{\log_{\alpha} \kappa}
         \text{.}
   \end{equation}
   \end{property}
  \begin{example}
    For the frequency decade band $\decade(f) \isdef \ClosedInterval{f, 10 f}$, a logarithmic frequency grid contains $F_{\decade} = \floor{\log_{\alpha} 10}$ excited lines.
  \end{example}
  \begin{example}
    The~\citep{HP3562A} has a logarithmic measurement mode that ensures that each decade is covered by $F_{\decade} = 80$ lines~ to measure an \gls{FRF}.
    This is equivalent to a grid spacing $\alpha = \sqrt[80]{10} \approx 1.029$.
  \end{example}

@@

  Using this extension, the frequency range covered by $f_k$ can then be calculated as
    \begin{align}
      \fMin{k} &= \mean{f_{k-1}}{f_k}\\
      \fMax{k} &= \mean{f_{k+1}}{f_k}
    \label{eq:fRange}
    \text{.}
    \end{align}
  This range can be seen as the power-spectral density counterpart of the discrete spectrum relation \eqref{eq:nearestLines}.

  Note that instead of the geometric mean, the arithmetic mean can be used in \eqref{eq:fRange} to calculate almost identical boundaries.

@@

  The obtained FRF is then fitted to the transfer function model \eqref{eq:secondOrderSystemDTidentifiable} using the \texttt{elis} function of the \FDIDENT~\citep{FDIDENT} for \MATLAB.
  This is a maximum likelihood estimator that fits a parametric model to the measured data in the frequency domain.
  \TODO{add formulas with cost function etc.}
  To aid convergence and avoid local minima, the initial values for the model are chosen equal to the parameters of the true system $G_0(z)$, so that we can focus completely on the variance of the estimations.

@@

  \begin{remark}
  All simulations  have been repeated fitted with the \MATLAB System Identification Toolbox by means of the \code{oe} function~\citep{TDIDENT}.
  The observed results were identical and hence are not elaborated further in this text.
  \end{remark}

@@

  The resulting frequency response function $\hat{G}$ for one such system is shown in \figref{fig:FRF1} together with the standard deviation $\sigma_{\hat{G}}$ of the estimated parametric model for each of the different excitation signals.
  The bias is at least $20\unit{dB}$ smaller than the standard  deviation.
  Therefore the \gls{MSE} will be determined mainly by the variance of the estimate instead of the bias.

  \begin{figure}[th]
    \centering
    \setlength\figureheight{5cm}
    \setlength\figurewidth{0.8\columnwidth}
    \input{\thisDir/figs/Sys02-4.tikz}
    \caption[Simulated FRF and its variance for different excitation signals.]{FRF of a second order system with $\fR=40\unit{Hz}$ and $\damping=0.2$
             along with the corresponding variance and bias of the parametric models obtained using three different kinds of multisine excitations.}%
    \label{fig:FRF1}
  \end{figure}

@@

\begin{subappendices}
  \section{Justification for considering only poles}
  \TODO{benadering van 1/2 systeem a.d.h.v. 0/2 systeem wanneer $\omega_z << \wn$ (of $\omega_z >> \wn)$: rond de resonantie zijn beiden ongeveer gelijk. Afgeleide naar de nul zijn klein rond de resonantie. Informatiematrix is langs twee kanten gewogen met die afgeleide, dus die nul maakt weinig uit.}
\end{subappendices}

### 2015-11-16 ch02-qlogms/sysid.tex
Quite some changes to CH02

The information matrix tells us how good the individual parameter estimates are. However, if we are interested in estimating a transfer function $G(s,\theta)$ the quality of this estimate is of more importance. It is possible to find an expression for the variance of $\hat{G}(j\omega,\theta)$ using the Gauss approximation formula \citep{Ljung1999, Pintelon2001}. This gives
\begin{align}
\var{\hat{G}(j\omega)} &= \frac{\partial G^H(j\omega)}{\partial \theta}\IF^{-1}\frac{\partial G(j\omega)}{\partial \theta}.
\label{eq:varG}
\end{align}

@@

 \subsection{A second order system}
Consider the second order system
\begin{align}
y_t &= G(s,\theta)=\frac{1}{\frac{s^2}{\wn^2} + 2\frac{\damping}{\wn}s + 1}u_t,
\label{eq:2ndOrderSys}
\end{align}
with parameters $\theta=[\wn\;\damping]^{\T}$. The resonance frequency of the system is $\wR \isdef \sqrt{1-\damping^2}\wn$
and the $3\unit{dB}$ bandwidth of the system is $\wdB \isdef 2\damping\wR$.
Since information is local in the frequency domain, the input signal power inside $\wdB$, contributes significantly to the information matrix while input power outside $\wdB$ has a smaller contribution. This can be seen by considering the optimal design for a second order system. From \citep{Goodwin1977} we have that the D-optimal design for \eqref{eq:2ndOrderSys} is a single sinusoid at frequency
\begin{align}
\omega^\star =\frac%
{\wn
  \sqrt{
    \left(1-2\damping^2\right)
    +
    \left(\left(1-2\damping^2\right)^2+15\right)^{1/2}
    }
  }
{\sqrt{5}}.
\label{eq:optFreq}
\end{align}
For systems with low damping, i.e $\damping\ll1$ , $\omega^\star\approx \wn$, which is inside the bandwidth of the system. 
From \eqref{eq:optFreq} it is clear that the optimal design depends on the resonance of the system to be identified. 
Input design can be based on a guess of $\wn$, however if this guess is wrong, the resulting estimates may be unacceptable~\citep{Rojas2007}.

When no information about the system is available, we propose using an input power spectrum with a logarithmic power distribution. This can be achieved either with a continuous bandlimited $1/f$ spectrum or a quasi-logarithmic multisine spectrum \citep{Pintelon2001}. We choose to work with an equidistant frequency grid with resolution $f_0$ and design the input in the frequency band $[\wn \kMin\; \wn \kMax]$, where the resolution is $k_0=1/T_{meas}$ and $T_{meas}$ is the measurement time. Hence, a finite measurement time limits the frequency resolution and therefore we have to put a lower limit on the frequency band of interest. Moreover, we can only impose the logarithmic distribution sufficiently well if $\kMin$ is large enough.

@@

Second, we consider the power spectrum of the quasi-logarithmic multisine signal [\cite{Pintelon2001}] given by
\begin{align}
  \phi_u(\omega) &= \frac{1}{M}\sum_{k=1}^{M}{\delta(\omega,\wn[K])},
\end{align}
where $\omega_{k}/\omega_{k+1} \approx \alpha_{f} = e^{\Delta}$.
Hence, the number of spectral lines inside \wdB is 
\begin{align}
\left\lfloor\ln{\left[\frac{1+\damping}{1-\damping}\right]}\Delta^{-1}\right\rfloor,
\end{align}
where $\lfloor\cdot\rfloor$ is the nearest integer rounded towards $0$. The number of spectral lines is independent of $\wn$.

@@

To construct a quasi-logarithmic multisine, only frequency lines $f_j$ within the set of available sine components $\left\{f_0, 2 f_0, \ldots, F f_0  \right\}$ are excited for which the ratio of two consecutive frequencies $f_j$ and $f_{j-1}$ is approximately constant: $\frac{f_{j-1}}{f_j} \approx \alpha_f$.

@@

For the quasi-log excitation, a frequency ratio of $\alpha_f = 1.05$ was chosen
and the amplitude spectrum was kept constant $A_k = A_{\mathrm{in}}$. This
amplitude was normalized such that the root mean square (RMS) value of $u[n]$
was equal to $1$.

The desired frequency ratio $\alpha_f$ may not be attainable for low frequencies,  due to a lack of frequency resolution.
This causes a less dense coverage of the lowest frequency bands.

>>>>

The information matrix tells us how good the individual parameter estimates are. However, if we are interested in estimating a transfer function $G(s,\theta)$ the quality of this estimate is of more importance. It is possible to find an expression for the variance of $\hat{G}(j\omega,\theta)$ using the Gauss approximation formula \citep{Ljung1999, Pintelon2012}. This gives
\begin{align}
\var{\hat{G}(j\omega)} &= \frac{\partial G^H(j\omega)}{\partial \theta}\IF^{-1}\frac{\partial G(j\omega)}{\partial \theta}.
\label{eq:varG}
\end{align}

@@

 \subsection{A second order system}
Consider the second order system
\begin{align}
y_t &= G(s,\theta) u_t = \frac{1}{\frac{s^2}{\wn^2} + 2\frac{\damping}{\wn}s + 1}u_t,
\label{eq:2ndOrderSys}
\end{align}
with parameters $\theta=[\wn\;\damping]^{\T}$. The resonance frequency of the system is $\wR \isdef \sqrt{1-\damping^2}\wn$
and the $3\unit{dB}$ bandwidth of the system is $\wdB \isdef 2\damping\wn$.
Since information is local in the frequency domain, the input signal power inside $\wdB$, contributes significantly to the information matrix while input power outside $\wdB$ has a smaller contribution. This can be seen by considering the optimal design for a second order system. From \citep[Example 6.4.5]{Goodwin1977} we have that the D-optimal design for \eqref{eq:2ndOrderSys} is a single sinusoid at frequency
\begin{align}
\omega^\star =\frac%
{\wn
  \sqrt{
    \left(1-2\damping^2\right)
    +
    \sqrt{\left(1-2\damping^2\right)^2+15}
    }
  }
{\sqrt{5}}.
\label{eq:optFreq}
\end{align}
For lightly-damped systems, i.e $\damping \ll 1$ , $\omega^\star\approx \wn$, which is inside the bandwidth of the system. 
From \eqref{eq:optFreq} it is clear that the optimal design depends on the resonance of the system to be identified. 
Input design can be based on a guess of $\wn$, however if this guess is wrong, the resulting estimates may be unacceptable~\citep{Rojas2007}.

When no information about the system is available, we propose using an input power spectrum with a logarithmic power distribution. This can be achieved either with a continuous bandlimited $1/f$ spectrum or a quasi-logarithmic multisine spectrum \citep{Pintelon2012}. 
We choose to work with an equidistant frequency grid with resolution $f_0$ and design the input in the frequency band $[\wn \kMin\; \wn \kMax]$, where the resolution is $k_0=1/\Tm$ and $\Tm$ is the measurement time. Hence, a finite measurement time limits the frequency resolution and therefore we have to put a lower limit on the frequency band of interest. Moreover, we can only impose the logarithmic distribution sufficiently well if $\kMin$ is large enough.

@@

Second, we consider the power spectrum of the quasi-logarithmic multisine signal [\cite{Pintelon2012}] given by
\begin{align}
  \phi_u(\omega) &= \frac{1}{M}\sum_{k=1}^{M}{\delta(\omega,\wn[K])},
\end{align}
where $\omega_{k}/\omega_{k+1} \approx \alpha = e^{\Delta}$.
Hence, the number of spectral lines inside \wdB is 
\begin{align}
  \floor{
    \log_{\alpha} \frac{1+\damping}{1-\damping} 
  }
  \text{,}
\end{align}
where $\floor{\placeholder}$ is the nearest integer rounded towards $0$. The number of spectral lines is independent of $\wn$.

@@

To construct a quasi-logarithmic multisine, only frequency lines $f_j$ within the set of available sine components $\left\{f_0, 2 f_0, \ldots, F f_0  \right\}$ are excited for which the ratio of two consecutive frequencies $f_j$ and $f_{j-1}$ is approximately constant: $\frac{f_{j-1}}{f_j} \approx \alpha$.

@@

For the quasi-log excitation, a frequency ratio of $\alpha = 1.05$ was chosen
and the amplitude spectrum was kept constant $A_k = A_{\mathrm{in}}$. This
amplitude was normalized such that the root mean square (RMS) value of $u[n]$
was equal to $1$.

The desired frequency ratio $\alpha$ may not be attainable for low frequencies,  due to a lack of frequency resolution.
This causes a less dense coverage of the lowest frequency bands.
